Biotic invasions fuck with the manageability of pollination networks

### 2017-01-25 paper/supp_info_2.Rmd
Script to calculate asymmetry matchings on the cluster

---
title: "Biotic invasions fuck with the manageability of pollination networks"
subtitle: "Supporting Information"
classoption: a4paper
csl: ecology-letters.csl
output:
  pdf_document:
    fig_crop: no
  html_document: default
  word_document: default
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \renewcommand{\figurename}{Supporting Figure}
documentclass: artikel1
bibliography: references.bib
---

@@

@@

prop <- expand.grid(c("asymmetry", "max_dep"), 
                    c("scaled_FALSE", "scaled_TRUE")) %>% plyr::alply(1)
driver <- c("../data/processed/matching_frequency") %>% 
  list.files(full.names = T) %>%
  lapply(list.files, full.names = T) %>% 
  unlist() %>% 
  lapply(read_allRDS)
driver <- mapply(function(x, y){
  attr(x, "type") <- as.character(y[,1])
  attr(x, "scaled") <- stringr::str_split(as.character(y[,2]), "_") %>% unlist() %>% extract(2) %>% as.logical()
  return(x)
}, driver, prop)
driver_bi <- "../data/processed/matching_frequency_bi" %>% 
  read_allRDS()
```

```{r plot theme, echo = FALSE, warning = F, message = F}
fer_theme <- theme_bw() +
	theme(text = element_text(family = "Times"),
	      title = element_text(size = 7.5, hjust = 0),
	      legend.title = element_text(size = 9),
	      legend.text = element_text(size = 8),
				axis.text = element_text(size = 8),
				axis.title = element_text(size = 9, hjust = 0.5), 
				strip.text = element_text(size = 9, hjust = 0), 
				strip.background = element_blank(),
				plot.margin = grid::unit(c(5, 0, 2, 0), "mm"))
```

@@

In the main text we show all maximum cardinality matchings for a example network in which the direction of control is based on the dependence asymmetries (Figure 1). For completeness, here we show all the maximum cardinality matchings for an example network in which the links are bidirectional (Supporting Figure 3).

Empirical network example here.

@@

>>>>

---
title: "Biotic invasions fuck with the manageability of pollination networks"
subtitle: "Supporting Information"
classoption: a4paper
csl: ecology-letters.csl
output:
  pdf_document:
    fig_crop: no
    keep_tex: yes
  html_document: default
  word_document: default
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{rotating}
- \renewcommand{\figurename}{Supporting Figure}
- \renewcommand{\tablename}{Supporting Table}
documentclass: artikel1
bibliography: references.bib
nocite: | 
  @Lopezaraiza-Mikel2007, @Bartomeus2008
---

@@

mutate_cond <- function(.data, condition, ..., envir = parent.frame()) {
  condition <- eval(substitute(condition), .data, envir)
  .data[condition, ] <- .data[condition, ] %>% dplyr::mutate(...)
  .data
}

@@

prop <- expand.grid(c("asymmetry", "max_dep"), 
                    c("scaled_FALSE", "scaled_TRUE")) %>% plyr::alply(1)
driver <- c("../data/processed/matching_frequency") %>% 
  list.files(full.names = T) %>%
  lapply(list.files, full.names = T) %>% 
  unlist() %>% 
  lapply(read_allRDS)
driver <- mapply(function(x, y){
  attr(x, "type") <- as.character(y[,1])
  attr(x, "scaled") <- stringr::str_split(as.character(y[,2]), "_") %>% unlist() %>% extract(2) %>% as.logical()
  return(x)
}, driver, prop)
driver_bi <- "../data/processed/matching_frequency_bi" %>% 
  read_allRDS()

n_matched <- readRDS("../data/processed/n_matched.rds")

```{r plot theme, echo = FALSE, warning = F, message = F}
fer_theme <- theme_bw() +
	theme(text = element_text(family = "Helvetica"),
	      title = element_text(size = 7, hjust = 0),
	      legend.title = element_text(size = 8),
	      legend.text = element_text(size = 7),
				axis.text = element_text(size = 7),
				axis.title = element_text(size = 8, hjust = 0.5), 
				strip.text = element_text(size = 8, hjust = 0), 
				strip.background = element_blank(),
				plot.margin = grid::unit(c(5, 0, 2, 0), "mm"))
```

@@

In the main text we show all maximum cardinality matchings for a simple example network. To further ilustrate our methodology, here we also show the approach for the smallest of our empirical networks, the uninvaded network at site 10 (Supporting Table 1; Supporting Figure 3). The largest component of this network is composed by 16 species of which 2 are non-invasive plants and the rest are pollinators. The one-to-one relationship between matched and superior nodes implies that in order to achieve full network controllability, most pollinators would be unmatched, and hence driver species that require external intervention. At the same time, both plants in the community, *Heracleum sphpndylum* and *Rubus fructicosus* and one of the pollinators *Orthotylus/Lygocorus* tend to be classified as superior nodes.

```{r fig-SA_exp, echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 6.2, results = "hide", fig.cap = "**Illustration of the procedure with one empirical network**. (a) The initial visitation network. (b) The directed network in which the direction of control and the magnitude of the asymmetry are determined based on the mutual dependences. (c) One of the possible maximum matchings calculated using the procedures illustrated in the Supporting Figure 1 and 2."}
par(mfrow=c(2,2), mar=c(0,0,0,0))
n <- keep_largest_component(net$SA_exp)
replace_names <- data.frame(sp_name = igraph::V(n)$name) %>%
	dplyr::mutate(sp_number = 1:nrow(.),
								sp_letter = letters[sp_number], 
								sp_name = stringr::str_replace(sp_name, "[0-9]", ""),
								sp_name = stringr::str_replace(sp_name, "\\?rutitarse", "rufitarse"))
set.seed(7)

la <- igraph::layout_nicely(n)
n %>%
	plot(vertex.color = v_c, 
			 vertex.size = 15,
			 vertex.label = replace_names$sp_letter,
			 vertex.label.family = fam,
			 vertex.label.color = l_c,
			 vertex.label.cex = 0.9,
			 edge.label = round(igraph::E(.)$weight,1), 
			 edge.label.family = fam,
			 edge.label.color = l_c,
			 edge.label.cex = l_s*0.8,
			 edge.curved = F,
			 edge.arrow.size = 0.5,
			 margin = ma, 
			 layout = la)
text(label_pos_x, label_pos_y, paste0(letters[1], ")"), adj = c(0,0))

as_net <- n %>%
	bipartite_digraph(type = "weight", keep = "all", weight.type = "asymmetry", scale = F)

as_net %>%	
	plot(vertex.color = v_c, 
			 vertex.size = 15,
			 vertex.label = replace_names$sp_letter,
			 vertex.label.family = fam,
			 vertex.label.color = l_c,
			 vertex.label.cex = 0.9,
			 edge.label = round(igraph::E(.)$weight,2), 
			 edge.label.family = fam,
			 edge.label.color = l_c,
			 edge.label.cex = l_s*0.8,
			 edge.curved = F,
			 edge.arrow.size = 0.5,
			 margin = ma, 
			 layout = la)

text(label_pos_x, label_pos_y, paste0(letters[2], ")"), adj = c(0,0))

base_as_net <- n
# E(as_net)$weight <- round(E(as_net)$weight, 1)

or <- as_all_mat %>%
	plyr::laply(function(x){
		sum(E(as_net)$weight[x])
	}) %>% order(decreasing = T)

as_all_mat <- as_all_mat[or]

1 %>%
	plyr::l_ply(function(x){

		E(as_net)$matched <- 2
		E(as_net)$matched[as_all_mat[[x]]] <- 4

		igraph::V(as_net)$matched <- 2
		igraph::V(as_net)$matched[igraph::V(as_net)$name %in%
																ends(as_net, as_all_mat[[x]])[, 2]] <- 4

		igraph::V(as_net)$superior <- 1
		igraph::V(as_net)$superior[igraph::V(as_net)$name %in%
																ends(as_net, as_all_mat[[x]])[, 1]] <- 2

		# names_fancy <- sp_names_fancy
		# names_fancy[igraph::V(as_net)$name %in%
		# 														ends(as_net, as_all_mat[[x]])[, 1]] <-
		# 	sp_names_fancy_b[igraph::V(as_net)$name %in%
		# 														ends(as_net, as_all_mat[[x]])[, 1]]
		# 
		matching_weight <- sum(E(as_net)$weight[as_all_mat[[x]]])

		as_net %>% 
			plot(layout = la, 
					 vertex.size = igraph::V(.)$superior * 7 + 7,
					 vertex.label = replace_names$sp_letter,
					 vertex.label.family = fam,
					 vertex.label.color = l_c,
					 vertex.label.font = igraph::V(.)$superior,
					 vertex.label.cex = igraph::V(.)$superior * 0.1 + 0.9,
					 # vertex.shape = igraph::V(.)$vertex.shape,
					 vertex.color = pal[igraph::V(.)$matched],
					 edge.color = pal[E(.)$matched],
					 edge.label = NA, 
					 edge.width = E(.)$matched/2,
					 edge.label.family = fam,
					 edge.label.color = igraph::E(.)$matched,
					 edge.label.cex = l_s,
					 edge.curved = F,
					 edge.arrow.size = 0.5,
					 margin = ma) %>% 
			print()

		igraph::V(as_net)$name <- sp_names

		text(label_pos_x, label_pos_y, paste0(letters[2+x], ")"), adj = c(0,0))
		text(text_pos_x + 0.2, text_pos_y + 0.1, cex = 0.95,
				 paste0("D = ", length(igraph::V(as_net)) - as_mat$matching_size, "\n",
				 			 "matching size: ", as_mat$matching_size, "\n",
				 			 "matching weight: ", round(matching_weight, 1)))
})

plot.new()

legend(0.28, 0.95, legend = replace_names$sp_name, horiz = F, bty = "n", text.font = 3, cex = 0.9)
legend(0.2, 0.95, legend = replace_names$sp_letter, horiz = F, bty = "n", text.font = 2, cex = 0.9)

# library(grid)
# library(gridExtra)
# 
# ttt <- tableGrob(replace_names) 
# grid.newpage()
# grid.draw(ttt)

In addition, here we also show all the maximum cardinality matchings for an example network in which the links are bidirectional (Supporting Figure 4). It is important to note that using bidirectional is problematic for a couple

@@

# S2: Properties of empirical networks

```{r calc-network-properties, echo = F, message = F}
net_prop <- net %>%
	plyr::ldply(function(x){
		x <- keep_largest_component(x)
		bipartite::networklevel(igraph::as_incidence_matrix(x, 
																												types = igraph::V(x)$type == "pla",
																												attr = "weight"),
														index = c("connectance", "web asymmetry", 
																			"ISA", 
																			"weighted NODF")) %>%
			as.list() %>% as.data.frame() %>%
			dplyr::mutate(n_sp = length(igraph::V(x)),
										n_pla = sum(igraph::V(x)$type == "pla"),
										n_pol = sum(igraph::V(x)$type == "pol"))
	})

library(xtable)

names(net_prop)[1] <- "net_name"

net_prop <- meta %>%
	dplyr::mutate(n_s_t = n_pla + n_pol) %>%
	dplyr::select(-n_pla, -n_pol) %>%
	dplyr::inner_join(net_prop)

The studied networks had species richess ranging between `r min(net_prop$n_s_t)` and `r max(net_prop$n_s_t)` (`r min(net_prop$n_sp)`-`r max(net_prop$n_sp)` when considering only the largest component in each network). As evidenced by the overall network asymmetry *AS* [@Bluthgen2007] networks had a low ratio of plants to pollinators. Furthermore, networks had relatively low levels of nestedness [when measured using the quantitative version of the NODF index; @Almeida-Neto2011]. Details for each network can be found in Supporting Table 1.

```{r table-network-properties, results = "asis", echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 4.8, sanitize = FALSE}

net_prop %>%
	dplyr::select(-net_name, -count, -net) %>%
	dplyr::filter(study != "ballantyne") %>%
	dplyr::arrange(study, site, inv) %>%
	mutate_cond(inv == FALSE, invader = "—") %>%
  dplyr::mutate(site = rep(1:dplyr::n_distinct(site), each = 2),
                inv = plyr::mapvalues(inv, c(TRUE, FALSE), c("yes", "no")),
                study = plyr::mapvalues(study, 
                                        c("bartomeus", "lopezaraiza"),
                                        c("Cap de Creus, Spain",
                                          "Bristol, United Kingdom"))) %>%
	dplyr::select(site, invader, n_s_t, n_sp, n_pla, n_pol, connectance, web.asymmetry, weighted.NODF, study) %>%
	dplyr::rename_("$R$" = "n_s_t",
								 "$n_p$" = "n_pla",
								 "$n_a$" = "n_pol",
								 "location" = "study",
								 "$n_s$" = "n_sp",
								 "$c$"= "connectance", 
								 "NODF" = "weighted.NODF",
								 # "$\\bar{a}$" = "interaction.strength.asymmetry",
								 "$AS$" = "web.asymmetry") %>%
	xtable::xtable(caption = "\\textbf{Properties of the analysed plant-pollinator communities}. Invasive plants were \\textit{Carpobrotus affine acinaciformis} (car), \\textit{Opuntia stricta} (op), and \\textit{Impatients grandulifera} (imp). All properties, with the exception of the networks' total species richness ($R$), correspond to the network's largest component. Specifically we show the number of species ($n_s$), the number of plants ($n_p$), the number of pollinators ($n_a$), the network connectance ($c$), the network assymetry ($AS$), and the network nestedness (NODF index). British networks were assembled by Lopezaraiza-Mikel et al. (2007), Spanish were networks assembled by Bartomeus et al. (2008).") %>%
	print(#floating.environment = "sidewaystable", 
		comment = FALSE, sanitize.text.function = function(x){x}, include.rownames = FALSE)

# knitr::kable(meta)
```

# S3: Visitation as a proxy for species interdependence

Visitation frequency has been shown to be an appropriate surrogate for inter-specific effects in pollination networks [@Vazquez2005; @Bascompte2006]. 
Nevertheless visitation is not equivalent to pollen deposition and might be insufficient to reflect the dependencies of plants on animals and vice versa [@Alarcon2010; @King2013]. 
We therefore investigated the effect of calculating the dependencies using visitation or pollination effectiveness and importance---two metrics more proximate to plant reproductive success (Supporting Figure 6).
We did this by comparing *(i)* the manageability of the community and *(ii)* the percentage of interactions that maintained the direction of dependency. 
To do that, we used data collected by @Ballantyne2015 from a low diversity pollination community at a dry lowland heathland in Dorset, UK  (50° 43.7'N 2° 07.2'W). 
First, deposition networks were quantified using the mean Single Visit Deposition---the number of conspecific pollen grains effectively deposited on a virgin stigma during a single visit by a particular animal [@NeEman2010; @King2013; @Ballantyne2015]. 
Second, visitation networks were constructed counting the visits to flowers during Single Visit Depositions. 
Finally, pollinator importance networks were constructed as the product of pollinator efficiency and visit frequency.

```{r fig-visitation-vs-deposition-vs-importance, echo = F, message = F, warning = F, fig.width = 6.2, fig.height = 4, results = "hide", fig.cap= "**Distribution of interaction weights for the pollen deposition, visitation and pollinator importance networks**. Note that the *x* axis in the density plots have been log-transformed."}
n_driver <- meta %>%
  dplyr::inner_join(n_matched) %>% 
  dplyr::mutate(n_driver = 1 - n_matched / (n_pla + n_pol))

p <- net[meta$net_name[meta$study == "ballantyne" & meta$count != "binary"]] %>%
	plyr::ldply(function(x) {
		data.frame(vname = igraph::E(x) %>% attr("vnames"), 
							weight = igraph::E(x)$weight)
	}, .inform = T) %>%
	dplyr::rename_("net_name" = ".id") %>%
	dplyr::mutate(net_name_f = factor(net_name,  
																		levels = c( 'bal_eff','bal_vis','bal_imp'),
																		labels = c("a) Deposition", "b) Visitation", "c) Importance")), 
	              pla = stringr::str_split_fixed(vname, "\\|", 2)[, 1],
	              pol = stringr::str_split_fixed(vname, "\\|", 2)[, 2]) %>%
  plyr::dlply("net_name_f", function(x){
    p1 <- x %>%
      ggplot(aes(x = pol, y = pla)) +
      geom_tile(aes(fill = weight, width = 0.9, height = 0.9)) + 
      coord_fixed() +
      scale_fill_gradient(low = "#f0f0f0", high = "#525252") +
      ylab("plants") +
      xlab("pollinators") + 
      fer_theme +
      theme(axis.text = element_blank(), 
            axis.ticks = element_blank(), 
            legend.position = "right", 
            legend.title = element_blank(), 
            plot.margin = grid::unit(c(6,0,2,5), "mm"),
            panel.grid = element_blank(),
            legend.key.width = grid::unit(4, "mm"),
            legend.key.height = grid::unit(4, "mm"))
    p2 <- x %>% ggplot(aes(x = weight)) +
      scale_x_log10() +
      stat_density(aes(y =..density..), geom = "line") +
      xlab("link weight") + 
      ylab("density") +
      ggtitle(x$net_name_f[1]) +
      fer_theme +
      theme(plot.margin = grid::unit(c(2,2,0,0), "mm"))
    list(p1, p2)
  })

cowplot::plot_grid(p[[1]][[2]], p[[1]][[1]], p[[2]][[2]], p[[2]][[1]], p[[3]][[2]], p[[3]][[1]], ncol = 2, rel_widths = c(1.5,2))  
```

We first investigated the effects at a network scale. 
Despite marked differences in the distribution of weights of the three networks, the minimum number of driver species to control the whole community was consistent among the three different approaches (`r 1-round(n_driver$n_driver[n_driver$net_name == "bal_eff"], digits = 2)` for deposition, `r 1-round(n_driver$n_driver[n_driver$net_name == "bal_vis"], digits = 2)` for the visitation, and `r 1-round(n_driver$n_driver[n_driver$net_name == "bal_imp"], digits = 2)` for the pollinator importance network).

The choice of weighting used can also have an impact on the relative importance of species.
Therefore we calculated the frequency that each species is present in the possible sets of driver species under the three schemes. 
Although visitation and deposition produce strikingly different results, we found a very strong agreement between the order produced by visitation and importance (Supporting Table 2). 
Finally, we investigated whether the asymmetry of mutual dependency, which defines the direction of control, was consistent among the three possible weighting schemes. We found again that the direction of the dominant dependency was maintained was consistent for 95\% of the interactions weighted by visitation or importance, the two most appropriate metrics for pollinator and plant dependency (Supporting Table 2).

```{r table-correlations-visitation-vs-efficiency, echo = FALSE, warning = F, message = F, results = "asis"}
d_sp <- driver %>%
  plyr::ldply(function(x) {
    get_frequencies(x) %>%
      dplyr::mutate(type = attr(x, "type"),
                    scaled = attr(x, "scaled"))
  }) %>%
  dplyr::inner_join(meta)

correlations <- d_sp %>%
	dplyr::filter(study == "ballantyne",
								threshold == 0.5,
								type == "asymmetry",
								scaled == F) %>%
	dplyr::group_by(net_name, threshold, type, scaled) %>%
	dplyr::mutate(d_weight = d_weight/max(d_weight),
								d_rank = rank(-d_weight, ties.method = "average")) %>%
	dplyr::group_by() %>%
	dplyr::select(net_name, species, d_rank) %>%
	tidyr::spread(net_name, d_rank) %>%
  dplyr::select(-species) %>%
  as.matrix() %>%
  Hmisc::rcorr(type = "spearman")

my_paste <- function(x, y){
  if(is.na(y)) return ("-")
  b <- ""
  if(y < 0.05) b <- ""
  paste0(b, round(x, digits = 2), " (", fix_small_pvalue(y, 3), ")", b)
}

dir_net_b <- net[meta$net_name[meta$study == "ballantyne"]] %>%
  lapply(function(x){
    x %>% 
      bipartite_digraph(type = "weight", keep = "all", weight.type = "asymmetry") %>%
      igraph::ends(igraph::E(.)) %>% 
      apply(1, function(y) paste(y[1], y[2]))
  })

# same order as the correlations 
dir_net_b <- dir_net_b[colnames(correlations$P)]

lt <- combn(length(dir_net_b), 2) %>%
  apply(2, function(x){
    length(intersect(dir_net_b[[x[1]]], dir_net_b[[x[2]]])) /
      length(dir_net_b[[x[1]]])
  }) %>%
  magrittr::multiply_by(100) %>% round() %>% paste0("%")

nam <- c("unweighted","deposition", "importance", "visitation")

mapply(my_paste, correlations$r, correlations$P) %>%
  matrix(ncol = length(dir_net_b)) %>% magrittr::set_colnames(nam) %>%
  magrittr::set_rownames(nam) %>% `[<-`(lower.tri(.), lt) %>% 
  xtable::xtable(caption = "\\textbf{Correlation of $n_d$ among network weighting schemes}. Spearman correlation coeffcients (with p-value) of the relative importance of species and the percentage of interactions that share the direction of dependency obtained using the three weighting schemes and an unweighted scheme.") %>%
	print(comment = FALSE, include.rownames = FALSE)

Altogether, evidence supports the idea that visitation is a suitable metric to estimate the mutual dependency of species pairs. First, it is directly related to pollinator foraging. Second, it produces results consistent, at least within our controllability framework, with plant reproductive success (as estimated by the importance metric).
